Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals

### 2021-05-03 writeup_final/writeup_final.tex
glossa_draft

\paragraph{ }Existing research on the phonetic effects of mixed-language production has largely focused on a few pairs of phonologically related languages, and used temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, great variation in results (cf sec.\ref{asymmetries} ) suggests that data from a wider variety of populations, language pairs,  and phonetic features is required in order to make meaningful generalizations. There is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. Nevertheless, widespread multilingualism in this part of the world suggests that phonetic
%line--rephrase
 behavior in these populations can add to the understanding of such cross-language interactions, as they are more likely to reflect real-world experience with mixed-language processing.

\paragraph{} The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties (F1 and F2) of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: code-switching and cued picture-naming. The results demonstrate a dynamic effect of transfer on L2 vowels, but no difference in the extent of transfer between the two paradigms. Findings are discussed in light of recent proposals about asymmetries in transfer effects and the role of connected speech in introducing discursive factors to transfer studies.

@@

\paragraph{}Studies examining mixed-language use within an utterance mostly analyze just the switched (target) token. Thus, very few direct measurements of the duration of transfer effects are available. One study which measured this \cite{bullock2009trying} did not find any residual effects on the matrix language following a switch, suggesting that transfer during code-switching is localized.

@@

  \paragraph{}These results highlight that beyond interaction during the cognitive processing of sounds, phonetic transfer is ultimately a linguistic phenomenon, and thus subject to language-specific constraints. Therefore, to make meaningful generalizations, it is imperative to consider data from a variety of language pairs. Existing research on phonetic transfer largely revolves around a small set of phonologically related languages. Thus, the present study extends the scope of this research to a new set of languages-- Indian English and Bengali (an Indo-Aryan language).

@@

 \subsection{Asymmetries between sounds:}\label{asymmetry between sounds}

@@

Table \ref{table_f1_fixed_effects} summarizes the results of the fixed effects from the model for F1 at each point. As expected, we observe a significant effect of gender, such that male speakers have a lower mean F1 than female speakers. Vowel is not a significant predictor of F1. This seems surprising, given that the phonemic category is expected to systematically affect vowel quality. However, a boxplot visualization (Figure\ref{boxplot_F1}) shows that the participants in this study vary greatly both in the relative heights of the two categories and the extent of height difference between them. Note that these measures are only from the English (unilingual) condition, suggesting that the variability is a feature of the participants' IE vowel categories, independent of the effects of switching. The lack of any independent effect of Context on F1 suggests that using English words in mixed utterances does not lead to a uniform lowering/raising across vowels. Instead, the significant interaction between Context and Vowel indicates a lowering in \nt{2}, but not in \nt{\ae}, in the mixed-language condition. The results also reveal a significant effect of Task, such that the sentence task (code-switching) lowers both vowels. This effect is strongest at the beginning of the vowel, and disappears by 45\%. This is an unexpected result, since we have no principled reason to expect code-switching to induce an across-board lowering of vowels. As noted earlier, there is no significant interaction of Task with any of the other fixed factors. We used a series of ANOVAs to compare the above model with a null model that differs in the absence of the main factor which concerns our hypothesis-- Vowel*Context-- at each point in the vowel. We find that the effect is not significant at 5\%, and increases over the course of articulating the vowel. 
%ANOVA table add

@@

@@

Our third research question concerned the experimental paradigms. \citeA{olson2013bilingual} proposes that in both scripted and spontaneous code-switching paradigms, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", and thus proposes cued picture-naming as a preferable alternative. This calls to question the comparability of results between studies using these different paradigms. We tested this by comparing productions of the same items from the same group of participants in a scripted CS paradigm and a cued picture-naming paradigm. We found no significant interaction of Task either with Context, or with the Vowel*Context interaction, suggesting that the extent of phonetic transfer was not affected by the paradigm. 
There are at least two plausible reasons for this-- (i) it is possible that the planning effects of CS do not translate to an increase/decrease in transfer at all. This would mean that the hyperarticulation seen, for example, in \citeA{muldner2019phonetics} is better interpreted as resulting from some other aspect of the experimental setup, rather than the act of code-switching itself. While planning could additionally affect the words preceding the switch, the effects on the switched token itself do not change. (ii) Additionally, our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, another possible explanation is that the discursive effects discussed by \citeA{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can shed light on this. The results of the present study suggest that for the purposes of observing dynamic phonetic transfer, elicitation through cued-picture-naming vs (controlled) code-switching does not seem to significantly alter the behavior under study, and thus the choice of paradigm can be based on other considerations. Moreover, this suggests that the results from such studies can be compared directly.

Paradigm did affect vowel quality, but this applied uniformly across experimental conditions-- all vowels were more backed and lowered in the code-switching task. Given that at least for F1, the effect is limited to the early stages of articulating the vowel, it is plausible that this is an artifact of the particular stimuli used here-- resulting from some corarticulatory effect that was absent in the picture-naming task. While this is intriguing, it is beyond the scope of the present study.

>>>>

\paragraph{ }Existing research on the phonetic effects of mixed-language production has largely focused on a few pairs of phonologically related languages, and used temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, great variation in results (cf sec.\ref{asymmetries} ) suggests that data from a wider variety of
%elaborate on 'great variation'-- differences between language pairs, plus sociolinguistic situation (effects of language experience etc)-- clearer. 'Types of features'--separate line?- Brief mention - discursive factors in VOT. 
 populations, language pairs,  and phonetic features is required in order to make meaningful generalizations. There is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. Nevertheless, widespread multilingualism in this part of the world suggests that phonetic
%line--rephrase
 behavior in these populations can add to the understanding of such cross-language interactions, as they are more likely to reflect real-world experience with mixed-language processing.

%
\paragraph{} The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties (F1 and F2) of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: code-switching and cued picture-naming. The results demonstrate a dynamic effect of transfer on L2 vowels, but no difference in the extent of transfer between the two paradigms. Findings are discussed in light of recent proposals about asymmetries in transfer effects and the role of connected speech in introducing discursive factors to transfer studies.

@@

\paragraph{}Studies examining mixed-language use within an utterance mostly analyze only the switched (target) token. Thus, very few direct measurements of the duration of transfer effects are available. One study which measured this \cite{bullock2009trying} did not find any residual effects on the matrix language following a switch, suggesting that transfer during code-switching is localized.

@@

  \paragraph{}These results highlight that beyond interaction during the cognitive processing of sounds, phonetic transfer is ultimately a linguistic phenomenon, and thus subject to language-specific phonological constraints. Therefore, to make meaningful generalizations, it is imperative to consider data from a variety of language pairs. Existing research on phonetic transfer largely revolves around a small set of phonologically related languages. Thus, the present study extends the scope of this research to a new set of languages-- Indian English and Bengali (an Indo-Aryan language).

@@

 \subsection{Asymmetries between sounds}\label{asymmetry between sounds}

@@

Table \ref{table_f1_fixed_effects} summarizes the results of the fixed effects from the model for F1 at each point. As expected, we observe a significant effect of gender, such that male speakers have a lower mean F1 than female speakers. Vowel is not a significant predictor of F1. This seems surprising, given that the phonemic category is expected to systematically affect vowel quality. However, a boxplot visualization (Figure\ref{boxplot_F1}) shows that the participants in this study vary greatly both in the relative heights of the two categories and the extent of height difference between them. Note that these measures are only from the English (unilingual) condition, suggesting that the variability is a feature of the participants' IE vowel categories, independent of the effects of switching. The lack of any independent effect of Context on F1 suggests that using English words in mixed utterances does not lead to a uniform lowering/raising across vowels. Instead, the significant interaction between Context and Vowel indicates a lowering in \nt{2}, but not in \nt{\ae}, in the mixed-language condition. The results also reveal a significant effect of Task, such that the sentence task (code-switching) leads to higher F1 values in both vowels, relative to the cued picture-naming task. This effect is strongest at the beginning of the vowel, and disappears by 45\%. This is an unexpected result, since we have no principled reason to expect code-switching to induce an across-board lowering of vowels. As noted earlier, there is no significant interaction of Task with any of the other fixed factors. We used a series of ANOVAs to compare the above model with a null model that differs in the absence of the main factor which concerns our hypothesis-- Vowel*Context-- at each point in the vowel. We find that the effect is not significant at 5\%, and increases over the course of articulating the vowel. 
%ANOVA table add

@@

% only ae -- shift significant?
% difference between vowels-- transfer--makes reference to phonological categories--
% knitr -- put the R code for figures in here so changes update here
%in intro--say-- paper argues that these asymetries (seen in previous lit) point to the (urgent:)) need for diversifying the languages and sound features studied in order to make generalizations
\begin{figure}

@@

Our third research question concerned the experimental paradigms. \citeA{olson2013bilingual} proposes that in both scripted and spontaneous code-switching paradigms, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", and thus proposes cued picture-naming as a preferable alternative. This calls to question the comparability of results between studies using these different paradigms. We tested this by comparing productions of the same items from the same group of participants in a scripted CS paradigm and a cued picture-naming paradigm. We found no significant interaction of Task either with Context, or with the Vowel*Context interaction, suggesting that the extent of phonetic transfer was not affected by the paradigm. 
There are at least two plausible reasons for this-- (i) it is possible that the planning effects of CS do not translate to an increase/decrease in observable transfer at all. This would mean that the hyperarticulation seen, for example, in \citeA{muldner2019phonetics} is better interpreted as resulting from some other aspect of the experimental setup, rather than the act of code-switching itself. While planning could additionally affect the words preceding the switch, the effects on the switched token itself do not change. (ii) Our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, another possible explanation is that the discursive effects discussed by \citeA{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can shed light on this. The results of the present study suggest that for the purposes of observing dynamic phonetic transfer, elicitation through cued-picture-naming vs (controlled) code-switching does not seem to significantly alter the behavior under study, and thus the choice of paradigm can be based on other considerations. Moreover, this suggests that (all else being equal) the results from such studies can be compared directly.

Paradigm did affect vowel quality, but this applied uniformly across experimental conditions-- all vowels were more backed and lowered in the code-switching task. Given that at least for F1, the effect is limited to the early stages of articulating the vowel, it is plausible that this lowering is an artifact of the particular stimuli used here-- resulting from some corarticulatory effect that was absent in the picture-naming task. While this is intriguing, it is beyond the scope of the present study.
